Linear response theory for core set identification

### 2013-02-20 writedown/linear.response/main.tex
also select Backbone

>>>>

\newpage
\section{Numerical results}
\subsection{Tilting double-well potential}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
  \caption{Two testing cases presented by this paper.
    Left: the double-well potential with tilting perturbation.
    Right: the single-well potential with splitting perturbation.}
  \label{fig:new-tmp1}
\end{figure}

We test the idea by a one-dimensional model system: one particle in a
tilting double-well potential. For convenience, we let the mass of the
particle to be 1 \textsf{amu}. The unperturbed
Hamiltonian of the system is given by:
\begin{align}
  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
\end{align}
with potential
\begin{align}
  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
\end{align}
Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
See the red line in Fig.~\ref{fig:tmp1}.
%$\textsf{kJ} / (\textsf{mol nm}^4)}$
The perturbation is given by
\begin{align}
  \vect C(\vect p, \vect q) = 0; \qquad
  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
\end{align}
Here $V(\vect q) = -\vect q$ is  effectively tilting the original
potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
by the following function:
\begin{align}
  F_e(t) = 
  \begin{cases}
    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
    F_e^{\textrm{max}} & t \geq t_c
  \end{cases}
\end{align}
$t_c = 20$~\textsf{ps} so that the perturbation increases slow
enough: the typical decaying time scale of the correlation function
in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
We choose $\tau = 1$~\textsf{ps}.
% We choose $\mt_\alpha$
% to be the propagator of the Langevin dynamics at temperature $150$
% \textsf{K} with $\alpha = 1$ \textsf{ps}.
Since the dynamics is in
1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
The larger
this value, the more stable the core sets are.

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-tilt-str2-simple.eps}
  \caption{
    Tilting double well potential testing case:
    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $F_e^{\textrm{max}} = 2.0$.
    From left to right:
    (a) $t=0$~\textsf{ps},
    the brutal force non-equilibirum simulation,
    (b) $t=20$~\textsf{ps},
    the brutal force non-equilibirum simulation,
    (c) $t=20$~\textsf{ps},
    response formula truncated at order $n=1$,
    and
    (d) $t=20$~\textsf{ps},
    response formula truncated at order $n=2$.
    The equilibirum state serves as the reference of
    the response simulations.
  }
  \label{fig:new-tmp2}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-tilt-str4-simple.eps}
  \caption{
    Tilting double well potential:
    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $F_e^{\textrm{max}} = 4.0$.
    From left to right:
    (a) $t=0$~\textsf{ps},
    the brutal force non-equilibirum simulation,
    (b) $t=20$~\textsf{ps},
    the brutal force non-equilibirum simulation,
    (c) $t=20$~\textsf{ps},
    response formula with reference process $F_e^{\max} = 0.0$
    (i.e. equilibirum simulation),
    and
    (d) $t=20$~\textsf{ps},
    response formula with reference process $F_e^{\max} = 2.0$.
    The response formula are truncated at order $n=1$.
  }
  \label{fig:new-tmp3}
\end{figure}

The Fig.~\ref{fig:new-tmp2} and \ref{fig:new-tmp3} present
the application of the response theory to
calculate core sets of the tilting double-well potential.
Fig.~\ref{fig:new-tmp2} presents
the snapshot at $t=20$~\textsf{ps}
of $\Delta\mt_\tau f(\vect x,t)$ under the maximum perturbation
strength $F_e^{\max} = 2.0$. Response formula truncated
at both order $n=1$ and $n=2$ are compared to 
the brutal force non-equilibirum simulation.
No obviously difference is shown. The second order response result
is slightly more precise than the first order one, but the
statistical error is also larger, which makes the profile noisy.
This figure tells us that
(1) the modeling error of the linear response (1st order) is
negligibly small; (2) the statistical error of numerically measuring the
linear response is also small.
Fig.~\ref{fig:new-tmp3} presents the $t=20$~\textsf{ps}
snapshots of $\Delta\mt_\tau f(\vect x,t)$
under a larger perturbations $F_e^{\max} = 4.0$.
The linear reponse using the equilibirum process as the reference
shows poor accuracy: the left well vanishes in this case,
while the linear response produces  artifical structures.
Since the profile is smooth, which indicates
the statistical error of calculating the linear term is small,
the deviation comes from the modellng error, namely the contribution
of higher order response terms.
By including the second order response term reduces the left
artifical core set, but also creates artifical finer grained
structures. The statistical accuracy is also worse than the linear response.
Plot (d) present the result of the linear response
formula using reference process $F_e^{\max} = 2.0$, which
is the process presented in Fig.~\ref{fig:new-tmp2}.
Much better modelling accuracy is achieved:
the left artifical core set disappear and the
statistical accuracy is also satisfactory.

% See Fig. \ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
% the direct non-equilibrium simulation and the response
% approximations are presented. Time slices $t = 0$, 10, 15 and 20
% \textsf{ps} are shown.  When the perturbation is small,
% i.e. $F_e^{\textrm{max}} = 2$, the direct non-equilibrium simulation
% and the linear response results are accurately consistent.
% The 2nd order response formula~\eqref{eqn:pert-approx-1} and
% \eqref{eqn:pert-approx-2} are even better, but the numerical error
% is bigger.
% As the
% perturbation grows stronger, the left well becomes less stable, while
% the right well becomes more stable.  When the perturbation is big,
% i.e. $F_e^{\textrm{max}} = 4$, the linear response result is no
% longer precise.
% When $t\geq 20$ \textsf{ps}, the left well basically
% disappears, however, the linear response calculation presents
% artifical occilation at the right well.
% The 2nd order response also has artifical effect, and the numerical
% uncertainty is high.
% The 4th rwo of Fig.~\ref{fig:tmp3} presents the 1st order result of the
% response formula~\eqref{eqn:pert-approx-1} and
% \eqref{eqn:pert-approx-2}, with
% respect to the reference state $F_e^{\textrm{max}} = 2$. The result
% is impressively improved compared with the linear response with respect to the
% equilibrium state.

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-split-str1-simple.eps}
  \caption{
    Splitting single-well potential testing case:
    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $F_e^{\textrm{max}} = 1.0$.
    From left to right:
    (a) $t=0$~\textsf{ps},
    the brutal force non-equilibirum simulation,
    (b) $t=20$~\textsf{ps},
    the brutal force non-equilibirum simulation,
    (c) $t=20$~\textsf{ps},
    response formula truncated at order $n=1$,
    and
    (d) $t=20$~\textsf{ps},
    response formula truncated at order $n=2$.
    The equilibirum state serves as the reference of
    the response simulations.
  }
  \label{fig:new-tmp5}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-split-str2-simple.eps}
  \caption{
    splitting single well potential:
    the plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $F_e^{\textrm{max}} = 2.0$.
    From left to right:
    (a) $t=0$~\textsf{ps},
    the brutal force non-equilibirum simulation,
    (b) $t=20$~\textsf{ps},
    the brutal force non-equilibirum simulation,
    (c) $t=20$~\textsf{ps},
    response formula with reference process $F_e^{\max} = 0.0$
    (i.e. equilibirum simulation),
    and
    (d) $t=20$~\textsf{ps},
    response formula with reference process $F_e^{\max} = 1.0$.
    The response formula are truncated at order $n=1$.
  }
  \label{fig:new-tmp6}
\end{figure}
